A preliminary constraint grammar for Russian

### 2015-03-30 2015-nodalida-ruscg/ruscg.tex
more stuffs
git-svn-id: https://svn.code.sf.net/p/apertium/svn/branches/papers/@59573 72bbbca6-d526-0410-a7d9-f06f51895060

Our purposes: Processing stressed wordforms and open-source machine translation

\cite{Karlsson-90}

@@

\section{Review of literature}

State-of-the-art morphological analysis in Russian is primarily based on
finite-state technology
\cite{Nozhov-03,Segalovich-03}.\footnote{Machine-learning approaches have also been successfully applied to Russian, most notably by 
\newcite{sharoff08lrec-mocky}.} Almost without exception, all large-scale 
morphological transducers of Russian are based on the
forward-looking \emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}.
This dictionary gives fine-grained morphological specifications for more than
100~000 words, including inflectional endings, morphophonemic alternations, 
stress patterns, exceptions, and idiosyncratic collocations.
We base our work on the recently developed [anonymous] morphological transducer, which is also based on Zaliznjak's dictionary (Author, 2015).\footnote{The [anonymous] transducer is implemented using a two-level 
morphology \cite{Koskenniemi-84}, and can be compiled using either {\tt xfst}
\cite{Beesley.Karttunen-03} or {\tt hfst} \cite{hfst-11}} This finite-state transducer (FST) generates all possible 
morphosyntactic readings of each wordform, regardless of its frequency or probability. Because Russian is a relatively highly inflected language, broad coverage is important, but widespread homonymy leads to the generation of many spurious readings, as discussed in Section~\ref{ambiguity} below. Because of this, one of the foundational steps in Russian natural language processing is homograph disambiguation.

@@

 -- and our Constraint 
Grammar\footnote{Implemented using vislcg3 constraint grammar parser
(\url{http://beta.visl.sdu.dk/cg3.html}).}
\cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
some readings based on syntactic context.

\cite{trosterud2009}

@@

% The preliminary tagging performance is P:,96.1%, R:,99.8% for POS tagging and P:,88.2%, R:,98.1% for complete morphosyntactic tagging.
\cite{peradin12}

@@

\section{Pipeline}

% FRAN

@@

\ex. \label{ex:FSToutput} 
	\a. \rus{новый}\texttt{{\small <adj><m><nn><sg><nom>}}
	\b. \rus{автомат}\texttt{{\small <n><m><nn><sg><nom>}}
	\c. 
% det/prn (e.g. pos... их, я, мой)
% 'over'generation (e.g. all dets get all gen/nbr/case combinations)
%^^^
%ROB

@@

\section{Development process}
%FRAN

@@

\section{Future work}

% ROB
% FRAN
% func analysis
% dep analysis

>>>>

This paper presents a preliminary constraint grammar for Russian. The principle 
objective of the constraint grammar is to produce a high recall grammar to serve
as input into other natural language processing tasks.

The paper is laid out as follows: section~\ref{sec:lit} presents a review of 
the literature on Russian language processing; section~\ref{ambiguity} gives
an overview of ambiguity in Russian; section~\ref{sec:pipeline} describes
our analysis pipeline; section~\ref{sec:devel} gives an account of our 
development process; section~\ref{sec:eval} presents an evaluation of the 
system, and sections~\ref{sec:future} and ~\ref{sec:conclusions} present future
work and conclusions.

%Our purposes: Processing stressed wordforms and open-source machine translation

%\cite{Karlsson-90}

@@

\section{Review of literature}
\label{sec:lit}
State-of-the-art morphological analysis in Russian is primarily based on
finite-state technology
\cite{Nozhov-03,Segalovich-03}.\footnote{Machine-learning approaches have also been successfully applied to Russian, most notably by 
\newcite{sharoff08lrec-mocky}.} Almost without exception, all large-scale 
morphological transducers of Russian are based on the
forward-looking \emph{Grammatical Dictionary of Russian} \cite{Zaliznjak-77}.
This dictionary gives fine-grained morphological specifications for more than
100~000 words, including inflectional endings, morphophonemic alternations, 
stress patterns, exceptions, and idiosyncratic collocations.
We base our work on the recently developed [anonymous] morphological transducer, which is also based on Zaliznjak's dictionary (Author, 2015).\footnote{The [anonymous] transducer is implemented using a two-level 
morphology \cite{Koskenniemi-84}, and can be compiled using either {\tt xfst}
\cite{Beesley.Karttunen-03} or {\tt hfst} \cite{hfst-11}} This finite-state transducer (FST) generates all possible 
morphosyntactic readings of each wordform, regardless of its frequency or probability. Because Russian is a relatively highly inflected language, broad coverage is important, but widespread homonymy leads to the generation of many spurious readings, as discussed in Section~\ref{ambiguity} below. Because of this, one of the foundational steps in Russian natural language processing is homograph disambiguation.

@@

% -- and our Constraint 
%Grammar\footnote{Implemented using vislcg3 constraint grammar parser
%(\url{http://beta.visl.sdu.dk/cg3.html}).}
%\cite{Karlsson-90,Karlsson.Voutilainen.ea-95} then removes
%some readings based on syntactic context.

%\cite{trosterud2009}

@@

% The preliminary tagging performance is P:,96.1%, R:,99.8% for POS tagging and P:,88.2%, R:,98.1% for complete morphosyntactic tagging.
% \cite{peradin12}

@@

\section{Pipeline}
\label{sec:pipeline}
% FRAN

@@

\ex. \label{ex:FSToutput} 
	\a. \rus{новый}\texttt{{\small <adj><m><nn><sg><nom>}}
	\b. \rus{автомат}\texttt{{\small <n><m><nn><sg><nom>}}
%	\c. 
% det/prn (e.g. pos... их, я, мой)
% 'over'generation (e.g. all dets get all gen/nbr/case combinations)
%^^^
%ROB

@@

\section{Development process}
\label{sec:devel}
%FRAN

@@

\section{Future work}
\label{sec:future}
% ROB
% FRAN
% func analysis
% dep analysis
